---
title: "Using Predictive Algorithms to Evaluate Human Exercise Quality"
author: "Ryan Hammer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold")
```
```{r load required packages}
library(ggplot2)
library(tidyverse)
library(caret)

```

## Introduction

The purpose of this analysis was to create a predictive model for evaluating the quality of exercise done by a person wearing a fitness based tracking device. The data had been split into a training set and a test set; links to each are given below:

Training: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
Test: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The original data source is: http://groupware.les.inf.puc-rio.br/har

## Data Summary
```{r load dataset, echo=FALSE, results='hold'}
url1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url1, "Training_Data.csv")
download.file(url2, "Testing_Data.csv")
train <- read.csv("Training_Data.csv") %>% tbl_df()
test <- read.csv("Testing_Data.csv") %>% tbl_df()
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
